---
title: "Problem Set 2"
subtitle: "Behavioral Economics, Boston College"
date: "Fall 2019"
author: "James Williams"
output:
  html_document:
    theme: lumen
    highlight: pygment
    number_sections: TRUE
---

<p style="color:red">
*The assignment is worth **100 points**. There are **26 questions**. You should have the following packages installed:*
</p>

```{r setup, results='hide', message=FALSE, warning=FALSE}
library(tidyverse)
library(cowplot)
library(lfe)
library(stargazer)
```

In this problem set you will summarize the paper ["Do Workers Work More if Wages Are High? Evidence from a Randomized Field Experiment"](https://www.aeaweb.org/articles?id=10.1257/aer.97.1.298) (Fehr and Goette, AER 2007) and recreate some of its findings. 


# Big picture

**1. What is the main question asked in this paper?**
The main question this paper attempts to answer is "Do workers work more when wages are high"?

**2. Recall the taxi cab studies where reference dependence is studied using observational data. What can an experimental study do that an observational study can't?**

An experimental study can directly examine any difference introduced in the form of an intervention to the groups it is analysing, the control group and th experimental group. Any difference in outcomes can be directly linked to the intervention introduced. In this example, by examining groups subject to a wage increase or not, the researchers are directly contribute any results that differ between the two groups to the increase in wage. 

**3. Summarize the field experiment design.** 

The feild experiment design examined messengers from two bike messaging company Veloblitz. The employees were split into two groups A and B based on a number assigned to them based on when they started working there. Then the memebers in the treatment group recieved a wage increase of roughly 25% for a 4 week period. Men's wages increased from .39 to .49/hr and women from .44 to .54/hr. Afterr the 4 week period ended for the intial treatment group, the initial control group, who recieved no increase in wage during the first 4 weeks, underwent the same wage increase the inital treatment group was subject to. This wage increase was incurred for a 4 week cycle as well. This was done as is pointed out on page 302 of the study so that "The point estimate of the treatment effect is completely independent of the heterogeniety between our subjects." in other words the study is able to paint a very acurate picture of the effects a temporary wage increase has on worker productivity. 

**4. Summarize the laboratory experiment design. Why was it included with the study?**
 The lab experiment took place 8 months after the inital feild experiment. The reserachers aimed to do this by examining choices made with a level of uncertainty. Two lotteries were presented to the members of the field experiment. The fist lottery had even odds (50%) to either win 8 CHF (Swiss franc), or lose 5 CHF. If subjects rejected this loterry they recieved 0 CHF. The next lottery was 6 independent completions of the first lottery stated above. Again, if this lottery was rejected 0CHF were recieved. Subjects were allowed to choose from, both, or neither of the lotteries. It was included in the study as stated on page 313 of the study to determine "whether the loss-averse messengers drive the negative effect of the wage increase on effort/shift or whether the messengers who are not loss averse drive this effect."
  
**5. Summarize the main results of the field experiment.**
When each group recieved the wage increase treatment the results were that workers in the treatment groups completed four more shifts than messengers in the control group, which was a significant finding. The control groups, however, earned roughly 5-6% more revenue per shift than the treatment groups.This implied that effort decreased with an increase in wage. There was also a finding that indicated that non-treated messebgers showed no statistical difference between the non treated messsengers and messengers from another bike company. 

**6. Summarize the main results of the laboratory experiment.**
42 messengers participated in the study forming Groups A and B. Of those 42 the results of the labratory experiment were:
  * 18 rejected both lotteries 
  * 8 rejected only lottery A 
  * 1 rejected only lottery B 
  * 14 accepted both
  
Rejecting both would show a higher rate of loss aversion than rejecting just A, and rejecting just A would show a higher rate of loss aversion than rejecting neither. They found that there was a 10% reduction in revenue per shift for those who rejected one or both lotteries , while there was not any for messengers who rejected neither lottery. 

**7. Why are these results valuable? What have we learned? Motivate your discussion with a real-world example.**
The results are consistent with refernece dependence loss aversion rather than the assumptions of the neoclassical model. This means that with a temporary increase in wages, workers will work more more hours. This comes with a caveat though, which is that because workers in the study exhibited loss aversion preneces, based on both the findings in the field and lab experiments, their marginal utility of earnings is diminishing after workers reach a certain threshold of income or a reference point. This causes workers to decrease effort but work more. 

A real world example similar to bike messengers is cab drivers. Cab drivers wage rates fluctuate day to day based on aseries of deman shocks such as weather, or holidays. Cab drivers themselves depict loss averse reference point dependent behavior, as they will make themselves more readily available on high demand days, until they hit their refence point, and feel like they do not have to work as much. 


# Theory

**Suppose the messenger's utility function is**

$$
v(e_t, x_t) = \gamma w_t e_t - g(e_t, x_t)
$$
**where $w_t$ is the wage rate in time $t$, $e_t$ is the messenger's effort, $\gamma$ is the marginal utility of lifetime wealth, and $g(e_t, x_t) = \frac{(\theta - x_t)e^2}{2}$ is the messenger's cost function constant disutility of effort $\theta$ and exogenous disutility shock $x_t \sim N(0,\sigma^2)$. Since $\mathbb{E}[x_t] = 0$, you can assume $x_t = 0$.**
$$
  v(e_t, x_t) = \gamma w_t e_t - \frac{(\theta - x_t)e^2}{2}
$$
**8. Show that the messenger chooses a level of effort so that the marginal benefit of working equals the marginal cost.**
when $x_t$ is assumed to be 0, the first order condition is 
$$
\frac{\partial U}{\partial e}=0= \gamma w_t - \theta e
$$


$$
\frac{\partial U}{\partial e}=\gamma w_t = \theta e
$$
By assuming $\theta$ is positive, the marginal benefet to working is equal to the margianl cost. The marginal benefit is the wage earned in a given shift multiplied by the utility of wealth. This captures the idea that working increases overall utility of wealth, as one is adding to their wealth with the wage earned. The second side of the equation shows the marginal cost, the undefined $\theta$ variable, multiplied by the effort devoted by the worker in a given shift. 



**9. Show that the messenger in equilibrium responds to higher wages with higher effort.**

$$
e^*= \frac{\gamma w_t}{\theta}
$$


$$ 
\frac{\partial e^*}{\partial w}= \frac{\gamma}{\theta}
$$


By assuming $\theta$ is positive, an increase in wages will lead to an increase in effort. This is because $\theta$ was assumed positive, and $\gamma$ is a positive variable. In other words we can say that a workers response in terms of 
effort is positive with an increase in wage presents itself, or that response is strictly greater than zero. 

$$
\frac{\partial e^*}{\partial w} >0
$$

**10. Write an R function that calculates $e_t^*$ for different levels of $w_t$. Set default values of $\theta=\gamma=1$. Then use `curve()` to plot the labor supply for $w_t \in [0,10]$.** 
```{r}
#w_t=x in this example
effort <- function(x,theta=1,gamma=1) {
  e=(gamma*x)/theta
  return(e)
}
effort(x=10)
curve(effort(x,theta=1,gamma=1), from=0, to=10)
```


**11. Now suppose utility is given by**

$$
v(e_t, x_t) = 
  \begin{cases}
    \gamma(w_t e_t - r) - g(e_t, x_t) &\quad \text{if} \quad w_t e_t \geq r \\[1em]
    \lambda\gamma(w_t e_t - r) - g(e_t, x_t) &\quad \text{if} \quad  w_t e_t < r \\
  \end{cases}
$$

**12. Show that how the messenger in equilibrium responds to higher wages depends on the reference point $r$. (Hint: recall there are three cases to consider.)**
$$
v(e_t, x_t) = 
  \begin{cases}
    \gamma(w_t e_t - r) - \frac{(\theta - x_t)e^2}{2}  &\quad \text{if} \quad w_t e_t \geq r \\[1em]
    \lambda\gamma(w_t e_t - r) - \frac{(\theta - x_t)e^2}{2} &\quad \text{if} \quad  w_t e_t < r \\
  \end{cases}
$$

When $x_t$ is assumed to be 0, the first order conditions are


$$
\frac{\partial U}{\partial e}=
\begin{cases}
    \gamma w_t - \theta e  &\quad \text{if} \quad w_t e_t \geq r \\[1em]
    \lambda\gamma w_t - \theta e &\quad \text{if} \quad  w_t e_t < r \\
  \end{cases}
$$

*Case 1: Below the reference point*


$$
\lambda\gamma w_t - \theta e <0
$$
$$
e^*= \frac {\lambda \gamma w_t}{\theta}
$$
$$
\frac {\partial e^*}{\partial w_t}= \frac {\lambda \gamma}{\theta} > 0

$$

If $\theta$ is assumed to be positive, the partial derivative of effort with respect to wage is strictly positive. Thus is becaus lambda is the loss averision variable, and is positive, $\gamma$ is utility of wealth, and is also positive. By assuming theta is positive we can see that a worker who is below their refernce point and is loss averse will work harder with an increase in wage, in order to reach the threshold of income, or their reference point, that they need to achieve in order to avoid incuring a monetarty loss. 

*Case 2: At the reference point*

$$
\lambda\gamma w_t - \theta e <0< \gamma w_t - \theta e 
$$

This is the case where the driver is content to only hit the refernce point. Because of this, the optimal level of effort the worker is willing to contribute is equal to i.e. :

$$
e^* = \frac {r}{w}
$$

because the refrenece point is contrained by the inequalities $w_te_t >= r$ and $w_t x e_t <r$. Therefore in the case where the worker is reference dependent :

$$
\frac {\partial e^*}{\partial h}= - \frac {r}{w^2}
$$

This is to say that the worker will work more on low demand days than on high demand days when the wage increases. Their response to the increase in wage is negative because on the high demand days when the wage increases, they would have to work less hours compared to low demand days when they would have to work more hours to reach the same refernece point to avoid monetary loss.  

*Case 3: Above the refence point* 

$$
\gamma w_t - \theta e >0
$$

$$
e^*= \frac {\gamma w_t}{\theta}
$$

$$
\frac {\partial e^*}{\partial w_t}= \frac {\gamma}{\theta} >0
$$

By assuming $\theta$ is positive, an increase in wages will lead to an increase in effort. This is because $\theta$ was assumed positive, and $\gamma$ is a positive variable. In other words we can say that a workers response in terms of effort is positive with an increase in wage presents itself, or that response is strictly greater than zero, which is consistent with the standard model. 

**13. Once more write an R function that calculates $e_t^*$ for different levels of $w_t$. Set default values of $\theta=\gamma=1$, $\lambda = 2$ and $r=3$. Then use `curve()` to plot the labor supply for $w_t \in [0,10]$.** 

```{r}
# x= w_t in the formula below
rd_effort <- function(x,r=3,theta=1, gamma=1, lambda=2){
  e=ifelse(x>=r, ((gamma*x)/theta), ((lambda*gamma*x)/theta))
  return(e)
}
curve(rd_effort(x,r=3,theta=1,gamma=1,lambda=2), from = 0, to= 10, main="Reference Dependence Effort") 
```



# Replication 

<p style="color:red">
*Use `theme_classic()` for all plots.*
<p style="color:red">

## Correlations in revenues across firms

<p style="color:red">
*For this section please use `dailycorrs.csv`.*
<p style="color:red">

**14. The authors show that earnings at Veloblitz and Flash are correlated. Show this with a scatter plot with a regression line and no confidence interval. Title your axes and the plot appropriately. Do not print the plot but assign it to an object called `p1`. **

```{r}
dailycorrs = read.csv("dailycorrs.csv")
head(dailycorrs, 10)
tail(dailycorrs, 10)
```
```{r}
p1<-
  qplot(logv, logf, data=dailycorrs, ylab = "Log of Flash Revenues/Week", xlab = "Log of Veloblitz Revenues/Week")+
  geom_smooth(method='lm', se= FALSE)+
  theme_classic()
```

**15. Next plot the kernel density estimates of revenues for both companies. Overlay the distributions and make the densities transparent so they are easily seen. Title your axes and the plot appropriately. Do not print the plot but assign it to an object called `p2`.**

```{r}
p2<-
  dailycorrs %>% 
  gather %>% 
  ggplot(aes(x=value, fill=key, color=key))+ 
  geom_density(alpha=0.25)+
  labs(x= "Distributions of Log of Revenues")+
  theme_classic()
  
 
```

**16. Now combine both plots using cowplot and label the plots with letters.**

```{r}
cowplot::plot_grid(p1,p2, labels = "auto")
```

## Tables 2 and 3

<p style="color:red">
*For this section please use `tables1to4.csv`.*
<p style="color:red">
```{r}
tables1to4= read.csv("tables1to4.csv")
head(tables1to4,10)
tail(tables1to4,10)
```


### Table 2

**On page 307 the authors write:**

>Table 2 controls for **individual fixed effects** by showing how, on average, the messengers' revenues deviate from their person-specific mean revenues. Thus, a positive number here indicates a positive deviation from the person-specific mean; a negative number indicates a negative deviation.

**17. Fixed effects are a way to control for *heterogeneity* across individuals that is *time invariant.* Why would we want to control for fixed effects? Give a reason how bike messengers could be different from each other, and how these differences might not vary over time.** 
When conducting econometric analysis one would want to control for fixed effects because these effects are constant across individuals, and thus have a hand in explaining any sort of relationship, correlation, or dependence you are seeking to unearth in analysis, when using data involovng individuals. Some examples for the bike messengers in the study at hand are daily diet, gender, age, weight, or height. In the case of bike messengers weight or daily diet may contribute to fatigue and ultimaltey end up influencing why a worker may or may not pick up more or less shifts. Because of that these effects must be incorporated in modeling. 

**18. Create a variable called `totrev_fe` and add it to the dataframe. This requires you to "average out" each individual's revenue for a block from their average revenue: $x_i^{fe} = x_{it} - \bar{x}_i$ where $x_i^{fe}$ is the fixed effect revenue for $i$.**

```{r}
tables1to4=
  tables1to4 %>% 
  group_by(fahrer) %>% 
  mutate(mean_totrev=mean(totrev)) %>% 
  mutate(totrev_fe=totrev-mean_totrev)
  
head(tables1to4)
```

**19. Use `summarise()` to recreate the findings in Table 2 for "Participating Messengers" using your new variable `totrev_fe`. (You do not have to calculate the differences in means.) In addition to calculating the fixed-effect controled means, calculate too the standard errors. Recall the standard error is $\frac{s_{jt}}{\sqrt{n_{jt}}}$ where $s_{jt}$ is the standard deviation for treatment $j$ in block $t$ and $n_{jt}$ are the corresponding number of observations. (Hint: use `n()` to count observations.) Each calculation should be named to a new variable. Assign the resulting dataframe to a new dataframe called `df_avg_revenue`.** 

```{r}
df_avg_revenue <- na.omit(tables1to4) %>%
  group_by(block, odd) %>%
    summarize(mean_revenues = mean(totrev_fe), standard_error = sd(totrev_fe)/n())
df_avg_revenue
```

**20. Plot `df_avg_revenue`. Use points for the means and error bars for standard errors of the means. Note the following:**

* To dodge the points and size them appropriately, use `geom_point(position=position_dodge(width=0.5), size=4)`
* To place the error bars use `geom_errorbar(aes(x=block, ymin = [MEAN] - [SE], ymax = [MEAN] + [SE]),width = .1,position=position_dodge(width=0.5))`
    + You need to replace `[MEAN]` with whatever you named your average revenues and `[SE]` with whatever you named your standard errors.
  
```{r}
ggplot(na.omit(df_avg_revenue), aes(block, mean_revenues, shape = as.factor(odd)))+
  geom_point(position = position_dodge(width = 0.5), size = 4)+
  geom_errorbar(aes(x=block, ymin = mean_revenues-standard_error, ymax = mean_revenues+standard_error), width = 0.1, position = position_dodge(width = 0.5))+
  theme_classic()
```

**21. Interpret the plot.**

The first block shows the average revenues of the two groups being very similar, before either group undergoes the treatment of the wage increase. The standard error bars overlap between the two groups, which indicates no. statistical signifignce in the differences between their mean revenues The second block depicts the Group A undergoing the wage increase, and as a result we see a great disparity between the average revenues between the two groups. Group A's mean revenues is approximatley 722, while Group B's is roughly -278. Intrestingly, and coenciding with expectations, we see the same disparity when the wage increase is applied to Group B in block 3, and not Group A. Group B's mean revenue is about 392, and Group A's is roughly -675. There is a large and signifigant difference in the mean revenues between the groups when the treatment of an increased wage is applied. 

### Table 3

**22. Recreate the point estimates in Model (1) in Table 3 by hand (you don't need to worry about the standard errors). Assign it to object `m1`. To recreate this model requires you to control for individual fixed effects and estimate the following equation:**

$$
y_{ijt} - \bar{y}_{ij} = \beta_1 (\text{H}_{ijt} - \bar{\text{H}}_{ij}) + \beta_2 (\text{B2}_{ijt} - \bar{\text{B2}}_{ij}) + \beta_3 (\text{B3}_{ijt} - \bar{\text{B3}}_{ij}) + (\varepsilon_{ijt} - \bar{\varepsilon}_{ij})
$$
#same coeffs amongst 3 models

**where $\text{H}$ is the variable `high`, $\text{B2}$ is the second block (`block == 2`) and $\text{B3}$ is the third block (`block == 3`).**

```{r}
m1.data<- tables1to4 %>% 
  filter(maxhigh ==1) %>%
  group_by(fahrer) %>%
  mutate(high_fe = high - mean(high)) %>%
  mutate(block2_fe = block2 - mean(block2)) %>%
  mutate(block3_fe = block3 - mean(block3))

m1<- lm(totrev_fe ~ high_fe + block2_fe + block3_fe, data = m1.data)
summary(m1) 
 
  
```

**23. Now recreate the same point estimates (ignoring the standard errors again) using `lm` and assign it to object `m2`. You are estimating**

$$
y_{ijt} - \beta_0 + \beta_1 \text{H}_{ijt} + \beta_2 \text{B2}_{ijt} + \beta_3 \text{B3}_{ijt} + \sum_{i=1}^{n} \alpha_i \text{F}_i + \varepsilon_{ijt}
$$
**where $\text{F}_i$ is the dummy variable for each messenger (`fahrer`).**
```{r}
m2 <- tables1to4 %>%
  filter(maxhigh == 1) %>%
  lm(totrev ~ high + block2 + block3 + as.factor(fahrer), data = .)
summary(m2)
```

**24. Now use the function [`felm()`](https://www.rdocumentation.org/packages/lfe/versions/2.8-3/topics/felm) from the `lfe` package to recreate Model (1), including the standard errors. Assign your estimates to the object `m3`. You are estimating**

$$
y_{ijt} = \alpha_i + \beta_1 \text{H}_{ijt} + \beta_2 \text{B2}_{ijt} + \beta_3 \text{B3}_{ijt} + \varepsilon_{ijt}
$$
**where $\alpha_i$ is the individual intercept (i.e. the individual fixed effect).**
**Note that the function call works as follows: `felm([y]~[x] | [grouping variable] | 0 | [clustering varaible], [name of data])`**

```{r}
m3 <- tables1to4 %>%
  filter(maxhigh ==1) %>%
  felm(totrev ~ high + block2 + block3 | fahrer | 0 | fahrer, data = .)
summary(m3)
```

**25. Compare the estimates in `m1`, `m2` and `m3`. What is the same? What is different? What would you say is the main advantage of using `felm()`?** 

M1 and m3 have the same coefficents, however one of the advantages of using the felm() command is that it provides an arguement to cluster the standard errors. This ensure more accuarte standard error results from unobserved correlations that may exist amongst the explanatory vairables, which could increase the likelihood of commiting a Type 1 error, and ultimately contibutes to more accurate slope coefficents being created for the model. The coeeficents for m2 are completely different from the identical models m1 and m3, because we have included a fixed effect in the form of including a dummy variable for every faher that participated in the study. The felm() functions main benefit is it reduces the amount of coding needed to create such a model. Controlling for the fixed effect using felm() is as simple as coding the variable we wish to control for between two | | symbols in the function. To create model 2 not using felm() we had to create de-meaned variables manually using dplyr(). Additionally, m3 does not contain an interncept unlike the first two models. 

**26. Recreate all of Table 3 and use [`stargazer()`](https://cran.r-project.org/web/packages/stargazer/vignettes/stargazer.pdf) to print the results.**
A few more models need to be created because we only made one needed in this table because m1=m3
```{r}
#All messengers in Veloblitz, treated and untreated
m4<-tables1to4 %>%
  filter(group != "Flash") %>%
  felm(totrev ~ high + block2 + block3 | fahrer | 0 | fahrer, data = .)
summary(m4)
```

```{r}
#Dummy for non-treated messangers
m5 <- felm(totrev ~ high + experiment + block2 + block3| fahrer | 0 | fahrer, data = tables1to4)
summary(m5)
```

```{r}
#Shifts/ 4 week period for messengers participating in study
m6 <- tables1to4 %>%
  filter(maxhigh ==1) %>%
  felm(shifts ~ high + block2 + block3 | fahrer | 0 | fahrer, data = .)
summary(m6)
```

```{r}
#Shifts/ 4 week period for all Veloblitz
m7 <- tables1to4 %>%
  filter(group != "Flash") %>%
  felm(shifts ~ high + block2 + block3 | fahrer | 0 | fahrer, data = .)
summary(m7)
```

```{r}
#Shifts/ 4 week period, all messengers
m8<- felm(shifts ~ high + experiment + block2 + block3 | fahrer | 0 | fahrer, data = tables1to4)
summary(m8)
```

```{r}
stargazer(m3,m4,m5,m6,m7,m8, dep.var.labels = c("Revenues per four-week period", "Shifts per four-week period"), covariate.labels = c("Treatment Dummy", "Dummy for nontreated at Veloblitz", "Treatment Period 1", "Treatment Period 2" ), align = TRUE, type = "html")
```



































